An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis

### 2020-08-04 Notes/Supplementary material/SupMat.tex
code

\title{Supplementary material for ``An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations''}

@@

The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
\begin{enumerate}[(i)]
\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
\end{enumerate}
Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules.  The detailed calculation of the derivation is not given here for the sake of conciseness. Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
\[
\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
\]
where $\bg_1$ and $\bg_2$ are defined by
\begin{align*}
\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
\end{align*}
First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors. Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$. Combining these two results leads to
\[
\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
\]
where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied. The delta method can therefore be applied; the result is as follows:
\begin{align*}
\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
\end{align*}
where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
\[
\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
\]

@@

\paragraph{When $n'\in I$}
From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$. Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here. Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian. The discrepancy should be made clear.]}. Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
\begin{equation}
\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
\label{eq:gamma.nn}
\end{equation}
when $K\to \infty$.
In addition, we remark that
\[
\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
\]
Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn}, we thus have
\begin{align*}
\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
\end{align*}
when $K\to\infty$.

@@

\section{Application to an Electrocardiogram}
We provide here an additional implementation of {\sf BoundEffRed}, applied to an electrocardiogram (ECG) dataset. The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $10$ segments of $50$ seconds each. Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (top panel), or EDMD (middle panel), or GPR (bottom panel). These extensions are superimposed to the ground-truth extension, plotted in red. The sharp and spiky ECG patterns make the AHM model too simplistic to describe this type of signal. Consequently, the forecast produced by {\sf SigExt} is moderately satisfactory.

\begin{figure}
\includegraphics[width=\textwidth]{ECGforecast.eps}
\caption{Extended ECG (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red).}
\label{fig:ecg}
\end{figure}

>>>>

\title{Supplementary material for ``An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis''}

@@

The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
\begin{enumerate}[(i)]
\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
\end{enumerate}
When these conditions are satisfied, the delta method gives the asymptotic behavior of $\bh^{(\ell)}=f^{(\ell)}(\bg)$, that is
\begin{equation}
\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
\label{eq:delta}
\end{equation}
where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
\[
\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
\]

Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules. {\color{red}Detail it}

Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
\[
\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
\]
where $\bg_1$ and $\bg_2$ are defined by
\begin{align*}
\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
\end{align*}
First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors. Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$. Hence, we have 
\[
\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
\]
where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$. Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied. The delta method can therefore be applied, and result~\eqref{eq:delta} is valid.

@@

\paragraph{When $n'\in I$}
From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$. Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here. Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian. The discrepancy should be made clear.]}. Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
\begin{equation}
\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
\label{eq:gamma.nn}
\end{equation}
when $K\to \infty$.
In addition, we remark that
\[
\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
\]
Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn} vanishes when $K\to\infty$. We thus have
\begin{align*}
\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
\end{align*}
when $K\to\infty$.

@@

\section{Application to an Electrocardiogram}
We provide here an additional implementation of {\sf BoundEffRed}, applied to an electrocardiogram (ECG) dataset. The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $10$ segments of $50$ seconds each. Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (first panel), or EDMD (second panel), GPR (third panel), or TBATS (fourth panel). These extensions are superimposed to the ground-truth extension, plotted in red. The sharp and spiky ECG patterns make the AHM model too simplistic to describe this type of signal. Consequently, the forecast produced by {\sf SigExt} is moderately satisfactory.

\begin{figure}
\includegraphics[width=\textwidth]{ECGforecast.eps}
\caption{Extended ECG (blue) obtained by the {\sf SigExt} forecasting (first panel), the EDMD forecasting (second panel), the GPR forecasting (third panel), and the TBATS forecasting (fourth panel), superimposed with the ground truth signal (dashed red).}
\label{fig:ecg}
\end{figure}

### 2020-08-04 Notes/results.tex
code

Corresponding results are given in Table~\ref{tab:mse.sine}. They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension. Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data. Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.

\begin{table}
\centering
\caption{AHM Signal: Performance of the Extension Methods}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
    &  & \\
   \hhline{|=#=|=|}
   {\sf SigExt} & $1.458$ & $1.069$ \\
   \hline
   Symmetric & $8.735$ & $0.001$ \\
   \hline
   EDMD & $2.313$ &$11.10$\\
   \hline
   GPR & $1.441$ &$244.7$ \\
   \hline
   TBATS & $1.501$ & $633.6$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

>>>>

The forecasting methods are applied to $1000$ realizations of the above-described signal. The mean and the standard deviation (SD) of the resulting MSEs are given in Table~\ref{tab:mse.sine}. They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension. Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data. Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.

\begin{table}
\centering
\caption{AHM Signal: Performance of the Extension Methods}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE (mean $\pm$ SD)}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
    &  & \\
   \hhline{|=#=|=|}
   {\sf SigExt} & $1.028 \pm  0.404$ & $1.069$ \\
   \hline
   Symmetric & $4.368 \pm  0.001$ & $0.001$ \\
   \hline
   EDMD & $1.156  \pm  0.003$ &$11.10$\\
   \hline
   GPR & $1.025  \pm  0.403$ &$244.7$ \\
   \hline
   TBATS & $1.501 \pm  1.354$ & $633.6$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}
